Details of specific functions

### 2023-11-17 inst/gitbook/03_specific-tutorials.Rmd
Updated and finalised test
good to go for merging!
Added vignette for multi.ace

 * `full` where the bootstrapping is entirely stochastic (*n* elements are replaced by any *m* elements drawn from the data)
 * `single` where only one random element is replaced by one other random element for each pseudo-replicate

@@

Level | Name | Description | Source | 
------|------|---------------------------------------------------|--------|
2 | `ancestral.dist` | The distance between an element and its ancestor | `dispRity` |
2 | `angles` | The angle of main variation of each dimensions | `dispRity` |
2 | `centroids`<sup>1</sup> | The distance between each element and the centroid of the ordinated space | `dispRity` |
1 | `convhull.surface` | The surface of the convex hull formed by all the elements | [`geometry`](https://cran.r-project.org/web/packages/geometry/index.html)`::convhulln$area` |
1 | `convhull.volume` | The volume of the convex hull formed by all the elements | [`geometry`](https://cran.r-project.org/web/packages/geometry/index.html)`::convhulln$vol` |
2 | `deviations` | The minimal distance between each element and a hyperplane | `dispRity` |
1 | `diagonal` | The longest distance in the ordinated space (like the diagonal in two dimensions) | `dispRity` |
1 | `disalignment` | The rejection of the centroid of a matrix from the major axis of another (typically an `"as.covar"` metric) | `dispRity` |
2 | `displacements` | The ratio between the distance from a reference and the distance from the centroid | `dispRity` |
1 | `edge.length.tree` | The edge lengths of the elements on a tree | `ape` |
1 | `ellipse.volume`<sup>1</sup> | The volume of the ellipsoid of the space | Donohue _et al._ (2013) |
1 | `func.div` | The functional divergence (the ratio of deviation from the centroid) | `dispRity` (similar to [`FD`](https://cran.r-project.org/web/packages/FD/index.html)`::dbFD$FDiv` but without abundance)|
1 | `func.eve` | The functional evenness (the minimal spanning tree distances evenness) | `dispRity` (similar to [`FD`](https://cran.r-project.org/web/packages/FD/index.html)`::dbFD$FEve` but without abundance)|
1 | `group.dist` | The distance between two groups | `dispRity` |
1 | `mode.val` | The modal value | `dispRity` |
1 | `n.ball.volume` | The hyper-spherical (*n*-ball) volume | `dispRity` |
2 | `neighbours` | The distance to specific neighbours (e.g. the nearest neighbours - by default) | `dispRity` |
2 | `pairwise.dist` | The pairwise distances between elements | [`vegan`](https://cran.r-project.org/web/packages/vegan/index.html)`::vegist` |
2 | `point.dist` | The distance between one group and the point of another group | `dispRity` |
2 | `projections` | The distance *on* (projection) or *from* (rejection) an arbitrary vector | `dispRity` |
1 | `projections.between` | `projections` metric applied between groups | `dispRity` |
2 | `projections.tree` | The `projections` metric but where the vector can be based on a tree | `dispRity` |
2 | `quantiles` | The *n*th quantile range per axis | `dispRity` |
2 | `radius` | The radius of each dimensions | `dispRity` |
2 | `ranges` | The range of each dimension | `dispRity` |
2 | `span.tree.length` | The minimal spanning tree length | [`vegan`](https://cran.r-project.org/web/packages/vegan/index.html)`::spantree` |
2 | `variances` | The variance of each dimension | `dispRity` |

@@

You can find more informations on the vast variety of metrics that you can use in your analysis in this [preprint](https://www.biorxiv.org/content/10.1101/801571v1).
<!-- biorXiv preprint -->

@@

@@

@@

>>>>

 * `"full"` where the bootstrapping is entirely stochastic (*n* elements are replaced by any *m* elements drawn from the data)
 * `"single"` where only one random element is replaced by one other random element for each pseudo-replicate
 * `"null"` where every element is resampled across the whole matrix (not just the subsets). I.e. for each subset of *n* elements, this algorithm resamples *n* elements across *ALL* subsets (not just the current one). If only one subset (or none) is used, this does the same as the `"full"` algorithm.

@@

Level | Name | Description | Source | 
------|------|---------------------------------------------------|--------|
2 | `ancestral.dist` | The distance between an element and its ancestor | `dispRity` |
2 | `angles` | The angle of main variation of each dimensions | `dispRity` |
2 | `centroids`<sup>1</sup> | The distance between each element and the centroid of the ordinated space | `dispRity` |
1 | `convhull.surface` | The surface of the convex hull formed by all the elements | [`geometry`](https://cran.r-project.org/web/packages/geometry/index.html)`::convhulln$area` |
1 | `convhull.volume` | The volume of the convex hull formed by all the elements | [`geometry`](https://cran.r-project.org/web/packages/geometry/index.html)`::convhulln$vol` |
2 | `deviations` | The minimal distance between each element and a hyperplane | `dispRity` |
1 | `diagonal` | The longest distance in the ordinated space (like the diagonal in two dimensions) | `dispRity` |
1 | `disalignment` | The rejection of the centroid of a matrix from the major axis of another (typically an `"as.covar"` metric) | `dispRity` |
2 | `displacements` | The ratio between the distance from a reference and the distance from the centroid | `dispRity` |
1 | `edge.length.tree` | The edge lengths of the elements on a tree | `ape` |
1 | `ellipse.volume`<sup>1</sup> | The volume of the ellipsoid of the space | Donohue _et al._ (2013) |
1 | `func.div` | The functional divergence (the ratio of deviation from the centroid) | `dispRity` (similar to [`FD`](https://cran.r-project.org/web/packages/FD/index.html)`::dbFD$FDiv` but without abundance)|
1 | `func.eve` | The functional evenness (the minimal spanning tree distances evenness) | `dispRity` (similar to [`FD`](https://cran.r-project.org/web/packages/FD/index.html)`::dbFD$FEve` but without abundance)|
1 | `group.dist` | The distance between two groups | `dispRity` |
1 | `mode.val` | The modal value | `dispRity` |
1 | `n.ball.volume` | The hyper-spherical (*n*-ball) volume | `dispRity` |
2 | `neighbours` | The distance to specific neighbours (e.g. the nearest neighbours - by default) | `dispRity` |
2 | `pairwise.dist` | The pairwise distances between elements | [`vegan`](https://cran.r-project.org/web/packages/vegan/index.html)`::vegist` |
2 | `point.dist` | The distance between one group and the point of another group | `dispRity` |
2 | `projections` | The distance *on* (projection) or *from* (rejection) an arbitrary vector | `dispRity` |
1 | `projections.between` | `projections` metric applied between groups | `dispRity` |
2 | `projections.tree` | The `projections` metric but where the vector can be based on a tree | `dispRity` |
2 | `quantiles` | The *n*th quantile range per axis | `dispRity` |
2 | `radius` | The radius of each dimensions | `dispRity` |
2 | `ranges` | The range of each dimension | `dispRity` |
1 | `roundness` | The integral of the ranked scaled eigenvalues of a variance-covariance matrix | `dispRity`
2 | `span.tree.length` | The minimal spanning tree length | [`vegan`](https://cran.r-project.org/web/packages/vegan/index.html)`::spantree` |
2 | `variances` | The variance of each dimension | `dispRity` |

@@

You can find more informations on the vast variety of metrics that you can use in your analysis in this [paper](https://onlinelibrary.wiley.com/doi/10.1002/ece3.6452).

@@

\begin{equation}
    roundness = \int_{i = 1}^{n}{\frac{\lambda_{i}}{\text{max}(\lambda)}}
\end{equation}

@@

#### Roundness

The roundness coefficient (or metric) ranges between 0 and 1 and expresses the distribution of and ellipse' major axis ranging from 1, a totally round ellipse (i.e. a circle) to 0 a totally flat ellipse (i.e. a line). A value of $0.5$ represents a regular ellipse where each major axis is half the size of the previous major axis. A value $> 0.5$ describes a pancake where the major axis distribution is convex (values close to 1 can be pictured in 3D as a cr\`{e}pes with the first two axis being rather big - a circle - and the third axis being particularly thin; values closer to $0.5$ can be pictured as flying saucers). Conversely, a value $< 0.5$ describes a cigar where the major axis distribution is concave (values close to 0 can be pictured in 3D as a spaghetti with the first axis rather big and the two next ones being small; values closer to $0.5$ can be pictured in 3D as a fat cigar).

This is what it looks for example for three simulated variance-covariance matrices in 3D:

```{r, eval = TRUE, echo = FALSE, message = FALSE}
## Note that the plotting functions requires the ellipse, rgl and VCVtools packages (but not need to load them)
require("ellipse")
require("rgl")
require("VCVtools")
## To install them
# install.packages(c("ellipse", "rgl"))
# devtools::install_github("TGuillerme/VCVtools")

## knitr RGL setup
# knitr::knit_hooks$set(webgl = hook_webgl)
options(rgl.useNULL = TRUE) # Suppress the separate window.

## A 3D sphere
sphere  <- VCVtools::make.VCV(shape = 1, dimensions = 3, min.thick = 0.01)
pancake <- VCVtools::make.VCV(shape = 0.75, dimensions = 3, min.thick = 0.01)
cigar   <- VCVtools::make.VCV(shape = 0.25, dimensions = 3, min.thick = 0.01)
```

```{r, fig.height = 4, fig.width = 4, echo = FALSE}
# options(rgl.printRglwidget = TRUE)
VCVtools::plot.VCV(sphere, dimensions = c(1,2,3), col.major.axes = c("orange", "blue", "darkgreen"), main = "Sphere's roundness: 1", lwd.major.axes = 3)
rglwidget()
```
```{r, fig.height = 4, fig.width = 4, echo = FALSE}
VCVtools::plot.VCV(pancake, dimensions = c(1,2,3), col.major.axes = c("orange", "blue", "darkgreen"), main = "Pancake's roundness: ~0.70", lwd.major.axes = 3)
rglwidget()
```
```{r, fig.height = 4, fig.width = 4, echo = FALSE}
VCVtools::plot.VCV(cigar, dimensions = c(1,2,3), col.major.axes = c("orange", "blue", "darkgreen"), main = "Cigar's roundness: ~0.25", lwd.major.axes = 3)
rglwidget()
```
```{r, eval = TRUE, fig.height = 4, fig.width = 12, echo = FALSE}
par(mfrow = c(1,3))
VCVtools::plot.roundness(sphere, dimensions = c(1,2,3), col.lines = c("orange", "blue", "darkgreen"), main = "Sphere's roundness: 1")
VCVtools::plot.roundness(pancake, dimensions = c(1,2,3), col.lines = c("orange", "blue", "darkgreen"), main = "Pancake's roundness: ~0.70")
VCVtools::plot.roundness(cigar, dimensions = c(1,2,3), col.lines = c("orange", "blue", "darkgreen"), main = "Cigar's roundness: ~0.25")
```

@@

Note that since the version `1.8` the trees and the matrices don't have to match allowing to run disparity analyses with variable matrices and trees.
This can be useful when running ancestral states estimations from a tree distribution where not all trees have the same topology.
